---
title: Data processing for the chemical denaturation of GFP
author: Erka Enkhbold
output:
  html_document: default
  pdf_document: default
---

Loading of data for absorbance at 395nm.

```{r}
abs_table <- read.csv("../data/exp_data.csv")
unfolded_abs <- mean(abs_table$Average[8:10]) # Assigning the average of 8x-10x as the absorbance of the unfolded state
native_abs <- abs_table$Average[1] # Assigning 1x as the absorbance of the native state
abs_table$fraction_unfolded <- 1 - (abs_table$Average - unfolded_abs)/(native_abs-unfolded_abs)

abs_table
```

Sample 7x seems off, so I will filter it out for analysis.
```{r}
processed_abs <- abs_table[c(1:6,8:10),c(1,2,6,7)]
processed_abs
```
Performing a sigmoidal fit on our experimental data using a non-linear least squares model.

```{r}
library(minpack.lm)

sigmoidal_fit <- nlsLM(
  fraction_unfolded ~ 1 / (1 + exp(-m * (Final_urea_concentration - dfifty))),
  data = processed_abs,
  start = list(m = 0.1, dfifty = median(processed_abs$Final_urea_concentration))
)


summary(sigmoidal_fit)
```
Visualising our experimental data

```{r}
library(ggplot2)

ggplot(data = processed_abs, aes(x = Final_urea_concentration, y = fraction_unfolded)) +
  geom_point(color = "chartreuse3", size = 3) +  # Scatter plot of data points
  theme_minimal()+
  theme(axis.line = element_line(color = "black",
                                 linewidth = 0.5,
                                 linetype = 1),
        
        )+
  
  geom_line(aes(y = predict(sigmoidal_fit)), color = "dark green", size = 1) +  # Fitted curve
  labs(
    x = "Urea Concentration (M)",
    y = "Fraction Unfolded"
  )+
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), expand = c(0,0))+
  scale_x_continuous(breaks = seq(0, 8, by = 0.8), expand = c(0,0))

  

```
Data processing for equillibrium constants and Gibbs Free Energy of each data point

```{r}
R <- 8.314  # J/(mol*K)
T <- 298.15 # Biochemical standard state (298.15 K)
abs_table$K <- abs_table$fraction_unfolded/(1-abs_table$fraction_unfolded)
abs_table$dGibbs <- -R*T*log(abs_table$K)
gibbs_processed <- subset(abs_table, !is.na(abs_table$dGibbs) & !is.infinite(abs_table$dGibbs)) #filtering NaN and infinite values
gibbs_processed
```
Linear regression of the Gibbs free energy change

```{r}

gibbs_lm <- lm(dGibbs~Final_urea_concentration, data = gibbs_processed)
summary(gibbs_lm)
```

Testing the assumption of normal distribution of errors
```{r}
hist(residuals(gibbs_lm))
```

Loading, wrangling and fitting reference data

```{r}
reference_data <- read.csv("../data/ref_data.csv")

ref_unfolded_abs <- mean(reference_data$Absorbance395[8:10]) # Assigning the average of 8x-10x as the absorbance of the unfolded state
ref_native_abs <- reference_data$Absorbance395[1] # Assigning 1x as the absorbance of the native state
reference_data$fraction_unfolded <- 1 - (reference_data$Absorbance395 - ref_unfolded_abs)/(ref_native_abs-ref_unfolded_abs)


ref_sigmoidal_fit <- nlsLM(
  fraction_unfolded ~ 1 / (1 + exp(-m * (Final_urea_concentration - dfifty))),
  data = reference_data,
  start = list(m = 0.1, dfifty = median(reference_data$Final_urea_concentration))
)
summary(ref_sigmoidal_fit)

```
Estimation of protein unfolding at 2, 4 and 6 M of urea.

```{r}
q2_concs <- c(2, 4, 6)

df_q2_concs <- data.frame(
  Final_urea_concentration = q2_concs
)

predict(ref_sigmoidal_fit, newdata = df_q2_concs) #Refernce dataset predictions

```
```{r}
predict(sigmoidal_fit, newdata = df_q2_concs) #Experimental dataset predictions
```


Visualising the reference data

```{r}
library(ggplot2)

ggplot(data = reference_data, aes(x = Final_urea_concentration, y = fraction_unfolded)) +
  geom_point(color = "chartreuse3", size = 3) +  # Scatter plot of data points
  theme_minimal()+
  theme(axis.line = element_line(color = "black",
                                 linewidth = 0.5,
                                 linetype = 1),
        
        )+
  
  geom_line(aes(y = predict(ref_sigmoidal_fit)), color = "dark green", size = 1) +  # Fitted curve
  labs(
    x = "Urea Concentration (M)",
    y = "Fraction Unfolded"
  )+
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), expand = c(0,0))+
  scale_x_continuous(breaks = seq(0, 8, by = 0.8), expand = c(0,0))
ggsave("../results/ref_denaturation_curve.png", height = 7.47, width = 8.69, units = "cm")

```
Processing reference data for the Gibbs Free Energy

```{r}
reference_data$K <- reference_data$fraction_unfolded/(1-reference_data$fraction_unfolded)
reference_data$dGibbs <- -R*T*log(reference_data$K)
ref_gibbs_processed <- subset(reference_data, !is.na(reference_data$dGibbs) & !is.infinite(reference_data$dGibbs))
ref_gibbs_processed
ref_gibbs_lm <- lm(dGibbs~Final_urea_concentration, data = ref_gibbs_processed)
summary(ref_gibbs_lm)
confint(ref_gibbs_lm)

```

Testing the assumption of normal distribution of errors
```{r}
hist(residuals(ref_gibbs_lm))
```


Visualising the Gibbs Free Energy distribution

```{r}
ggplot(data = ref_gibbs_processed) + aes(x = Final_urea_concentration, y = dGibbs) + geom_point() + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "dark green") + 
  theme_minimal()+
  theme(axis.line = element_line(color = "black",
                                 linewidth = 0.5,
                                 linetype = 1))+
  labs(x = "Urea Concentration (M)",y = "ΔG (J mol-1)")+
  scale_x_continuous(limits=c(0, 8), breaks=seq(0,8,0.8))+
  scale_y_continuous(limits=c(-15000, 15000), breaks=seq(-15000, 15000, 5000))+
  coord_cartesian(expand = FALSE)  

ggsave("../results/ref_gibbs.png", height = 7.47, width = 8.69, units = "cm")
```

The linear model calculates a highly significant (p<0.001) intercept of 11968 J mol-1, indicating a valid estimate of 12kJ mol-1 or 2.9 kcal mol-1 for the Gibbs Free Energy change associated with GFP folding.






